Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-11-08 MDPI_submission.Rmd
Commit 2023-11-08 15:20

abstract: |
  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
  Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales.
  Periods of positive trends are referred as brightening periods and of negative trends as dimming periods.
  Here we studied 30 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
  The clear-sky data were identified by applying a cloud screening algorithm. We found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $0.35\,\%/\text{year}$ for clear-sky conditions, and $-0.28\,\%/\text{year}$ for cloudy conditions.
  We have also investigated the consistency of these trends, the effect of the solar zenith angle, and their seasonal variability.
  Finally, we identified some anomalies in the long term SDR trends for all sky conditions.

# back matter
keywords: |
  GHI; SDR; Solar Brigthening/Dimming, aerosols, clouds.
## list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline

@@

```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky
    "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky
    "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky
```

@@

```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky
    "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky
    "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky
```

@@

```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA before and after noon. The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\\circ$ for each day, before and after local noon.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics("./images/SzaTrends-1.png")
knitr::include_graphics("./images/SzaTrends-4.png" )
knitr::include_graphics("./images/SzaTrends-7.png")
```

@@

## Long term trends by season of year

@@

```{r seasonalALL, echo=F, fig.cap="Linear trends (black lines) of monthly mean SDR season (columns of graphs) for the three sky conditions (rows of graphs). The vertical scale is common for all graphs.", out.width="100%"}
knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")
```

>>>>

abstract: |
  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
  Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales.
  Periods of positive trends are referred as brightening periods and of negative trends as dimming periods.
  Here we studied 30 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
  The clear-sky data were identified by applying a cloud screening algorithm. We found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $0.35\,\%/\text{year}$ for clear-sky conditions, and $-0.28\,\%/\text{year}$ for cloudy conditions.
  We have also investigated the consistency of these trends, the effect of the solar zenith angle, and their seasonal variability.
  Finally, we identified some anomalies in the long term SDR trends for all sky conditions by examining the cumulative sums of monthly anomalies from the climatological mean.

# back matter
keywords: |
  GHI; SDR; solar radiation; Solar Brigthening/Dimming; aerosols; clouds.
## list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline

@@

```{r cusummonth, echo = FALSE, fig.cap='Cumulative sum plots of the monthly SDR anomalies in (%) for different sky conditions.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky
    "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky
    "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky
```

@@

```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Cumulative sum plots of monthly SDR anomalies in (%) for different sky conditions after removing the long-term linear trend.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky
    "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky
    "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky
```

@@

```{r szatrends, echo = FALSE, fig.cap='Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics("./images/SzaTrends-1.png")
knitr::include_graphics("./images/SzaTrends-4.png")
knitr::include_graphics("./images/SzaTrends-7.png")
```

@@

## Long term trends by season

@@

```{r seasonalALL, echo=F, fig.cap="Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).", out.width="100%"}
knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")
```

### 2023-11-08 MDPI_submission.tex
Commit 2023-11-08 15:20

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The shortwave downward solar irradiance (SDR) is an important
factor that drives climate processes, production and can affect all
living organisms. Observations of SDR at different locations around the
world with different environmental characteristics have been used to
investigate its long-term variability and trends at different time
scales. Periods of positive trends are referred as brightening periods
and of negative trends as dimming periods. Here we studied 30 years of
pyranometer data in Thessaloniki, Greece, under three types of sky
conditions (clear sky, cloudy sky and all sky). The clear-sky data were
identified by applying a cloud screening algorithm. We found a positive
trend of \(0.38\,\%/\text{year}\) for all-sky, \(0.35\,\%/\text{year}\)
for clear-sky conditions, and \(-0.28\,\%/\text{year}\) for cloudy
conditions. We have also investigated the consistency of these trends,
the effect of the solar zenith angle, and their seasonal variability.
Finally, we identified some anomalies in the long term SDR trends for
all sky conditions.}

% Keywords
\keyword{GHI; SDR; Solar Brigthening/Dimming, aerosols, clouds.}

@@

\caption{Running cumulative sum plots of the monthly SDR.}\label{fig:cusummonth}
\end{figure}

@@

\caption{Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.}\label{fig:cusumnotrendmonthly}
\end{figure}

@@

\caption{Distribution of the SDR's long term trends by SZA before and after noon. The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon.}\label{fig:szatrends}
\end{figure}

\hypertarget{long-term-trends-by-season-of-year}{%
\subsection{Long term trends by season of
year}\label{long-term-trends-by-season-of-year}}

@@

\caption{Linear trends (black lines) of monthly mean SDR season (columns of graphs) for the three sky conditions (rows of graphs). The vertical scale is common for all graphs.}\label{fig:seasonalALL}
\end{figure}

>>>>

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The shortwave downward solar irradiance (SDR) is an important
factor that drives climate processes, production and can affect all
living organisms. Observations of SDR at different locations around the
world with different environmental characteristics have been used to
investigate its long-term variability and trends at different time
scales. Periods of positive trends are referred as brightening periods
and of negative trends as dimming periods. Here we studied 30 years of
pyranometer data in Thessaloniki, Greece, under three types of sky
conditions (clear sky, cloudy sky and all sky). The clear-sky data were
identified by applying a cloud screening algorithm. We found a positive
trend of \(0.38\,\%/\text{year}\) for all-sky, \(0.35\,\%/\text{year}\)
for clear-sky conditions, and \(-0.28\,\%/\text{year}\) for cloudy
conditions. We have also investigated the consistency of these trends,
the effect of the solar zenith angle, and their seasonal variability.
Finally, we identified some anomalies in the long term SDR trends for
all sky conditions by examining the cumulative sums of monthly anomalies
from the climatological mean.}

% Keywords
\keyword{GHI; SDR; solar radiation; Solar Brigthening/Dimming; aerosols;
clouds.}

@@

\caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions.}\label{fig:cusummonth}
\end{figure}

@@

\caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend.}\label{fig:cusumnotrendmonthly}
\end{figure}

@@

\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
\end{figure}

\hypertarget{long-term-trends-by-season}{%
\subsection{Long term trends by
season}\label{long-term-trends-by-season}}

@@

\caption{Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).}\label{fig:seasonalALL}
\end{figure}
